MG7: Configurable and scalable 16S data analysis

### 2016-07-25 docs/mg7-paper/out/paper.tex
update funding
update author contributions

\protect\hyperlink{statika}{Statika} is a Scala library developed by
\emph{AA} and \emph{EP-T} which serves as a way of defining and
composing machine behaviors statically. The main component are
\textbf{bundles}. Each bundle declares a sequence of computations (its
behavior) which will be executed in an \textbf{environment}. A bundle
can \emph{depend} on other bundles, and when being executed by an
environment, its DAG (Directed Acyclic Graph) of dependencies is
linearized and run in sequence. In our use, bundles correspond to what
an EC2 instance should do and an environment to an AMI (Amazon Machine
Image) which prepares the basic configuration, downloads the Scala code
and runs it.

@@

\protect\hyperlink{datasets}{Datasets} is a Scala library developed by
\emph{AA} and \emph{EP-T} with the goal of being a Scala-embedded
mini-language for datasets and their locations. \textbf{Data} is
represented as type-indexed fields: keys are modeled as singleton types,
and values correspond to what could be called a denotation of the key: a
value of type \texttt{Location} tagged with the key type. Then a
\textbf{Dataset} is essentially a collection of data, which are
guaranteed statically to be different through type-level predicates,
making use of the value--type correspondence which can be established
through singleton types and implicits. A dataset location is then just a
list of locations formed by locations of each dataset key. All this is
based on what could be described as an embedding in Scala of an
extensible record system with concatenation on disjoint labels, in the
spirit of \citep{harper1990extensible} \citep{harper1991record}. For
that \emph{Datasets} uses the {[}ohnosequences/cosas{]}{[}cosas{]}
library.

@@

\protect\hyperlink{loquat}{Loquat} is a library developed by \emph{AA},
\emph{EK} and \emph{EP-T} designed for the execution of embarrassingly
parallel tasks using S3, SQS and EC2 Amazon services.

@@

@@

\begin{itemize}
\tightlist
\item
  \textbf{AA} developed \emph{MG7}, \emph{Statika}, \emph{Datasets}, and
  \emph{aws-scala-tools}; wrote the paper;
\item
  \textbf{EK} developed \emph{nispero} (a prototype for \emph{Loquat}
  \citep{kovach2014nispero}) and \emph{aws-scala-tools}.
\item
  \textbf{MM} \emph{MG7} workflow design; curation and design of the
  \emph{16S-DB7} reference database; wrote the paper.
\item
  \textbf{PPT} curation, design, data extraction code of the
  \emph{16S-DB7} reference database.
\item
  \textbf{EP} \emph{MG7} workflow design; wrote the paper.
\item
  \textbf{RT} \emph{MG7} workflow design, assignment strategy; curation
  and design of the \emph{16S-DB7} reference database; wrote the paper.
\item
  \textbf{EPT} developed \emph{MG7}, \emph{Statika}, \emph{Datasets},
  \emph{FLASh/BLAST eDSLs}; data analysis approach and design; wrote the
  paper.
\end{itemize}

@@

\section{Acknowledgements}\label{acknowledgements}

\emph{Funding:} The two first authors are funded by INTERCROSSING (Grant
289974).

>>>>

\protect\hyperlink{statika}{Statika} is a Scala library developed by
\emph{AA} and \emph{EPT} which serves as a way of defining and composing
machine behaviors statically. The main component are \textbf{bundles}.
Each bundle declares a sequence of computations (its behavior) which
will be executed in an \textbf{environment}. A bundle can \emph{depend}
on other bundles, and when being executed by an environment, its DAG
(Directed Acyclic Graph) of dependencies is linearized and run in
sequence. In our use, bundles correspond to what an EC2 instance should
do and an environment to an AMI (Amazon Machine Image) which prepares
the basic configuration, downloads the Scala code and runs it.

@@

\protect\hyperlink{datasets}{Datasets} is a Scala library developed by
\emph{AA} and \emph{EPT} with the goal of being a Scala-embedded
mini-language for datasets and their locations. \textbf{Data} is
represented as type-indexed fields: keys are modeled as singleton types,
and values correspond to what could be called a denotation of the key: a
value of type \texttt{Location} tagged with the key type. Then a
\textbf{Dataset} is essentially a collection of data, which are
guaranteed statically to be different through type-level predicates,
making use of the value--type correspondence which can be established
through singleton types and implicits. A dataset location is then just a
list of locations formed by locations of each dataset key. All this is
based on what could be described as an embedding in Scala of an
extensible record system with concatenation on disjoint labels, in the
spirit of \citep{harper1990extensible} \citep{harper1991record}. For
that \emph{Datasets} uses the {[}ohnosequences/cosas{]}{[}cosas{]}
library.

@@

\protect\hyperlink{loquat}{Loquat} is a library developed by \emph{AA},
\emph{EK} and \emph{EPT} designed for the execution of embarrassingly
parallel tasks using S3, SQS and EC2 Amazon services.

@@

Partially funded by ITN INTERCROSSING (Grant 289974) and Cardiobiome
project ITC-20151148.

@@

\begin{itemize}
\tightlist
\item
  \textbf{AA} developed \emph{MG7}, \emph{Loquat}, \emph{Statika},
  \emph{Datasets}, and \emph{aws-scala-tools}; wrote the paper;
\item
  \textbf{EK} developed \emph{nispero} (a prototype for \emph{Loquat}
  \citep{kovach2014nispero}) and \emph{aws-scala-tools}.
\item
  \textbf{MM} \emph{MG7} workflow design; curation and design of the
  \emph{16S-DB7} reference database; wrote the paper.
\item
  \textbf{PPT} design and development of the first \emph{MG7} prototype
\item
  \textbf{EP} \emph{MG7} workflow design; wrote the paper.
\item
  \textbf{RT} \emph{MG7} workflow design, assignment strategy; curation
  and design of the \emph{16S-DB7} reference database; wrote the paper.
\item
  \textbf{EPT} developed \emph{MG7}, \emph{Statika}, \emph{Datasets},
  \emph{FLASh/BLAST eDSLs}; data analysis approach and design; reference
  database automated curation and filtering; wrote the paper.
\end{itemize}

@@
